---
title: "Thesis Data Aggregation/Cleaning"
author: "Yifei Chen"
date: "2024-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( 
  echo = TRUE, 
  message = FALSE,
  warning = FALSE
  )
```

```{r}
# Load necessary library
library(ggplot2)
library(tidyverse)
library (broom)
library(psych)
library(dplyr)
library(tidyr)
library(readr)
library(purrr)
library(tibble)
library(stringr)
library(ez)
library(multcomp)
library(jsonlite)
```

# Data Aggregation and Cleaning

## Preparation

Put all the csv files into a list and set the conditions appropriately.

```{r}
# List all CSV files in the directory
csv_files <- list.files(path = "Raw Data", pattern = "\\.csv$", full.names = TRUE)
```

Initialize an empty data frame to store the results.

```{r}
# Initialize an empty dataframe to store the results
GenRepRec_clean <- data.frame(stringsAsFactors = FALSE)
```

## Adding in participant ID

```{r}
# Function to process a single file
process_file_id <- function(file_name) {
  # Read the CSV file
  df <- read.csv(file_name, stringsAsFactors = FALSE)
  
  # Extract the string from the 'sona_id' column
  sona_id <- df$sona_id[1]
  
  return(as.data.frame(sona_id))
}

# Process each file and combine results
results_id <- map_df(csv_files, process_file_id)

# Rename the columns
colnames(results_id) <- c('Participant ID')

# Add the combined results into the data frames column-wise
GenRepRec_clean <- results_id
```

## Calculating the accuracy rate

### associative inference choice trials

Calculate the accuracy rate of associative inference choice trials (by trial type and session) across all csv files.

```{r}
# Function to process each CSV file
process_assinf_choice_acc <- function(file_path) {
  # Read the data
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Convert relevant columns to string
  data$assinfTrialType <- as.factor(data$assinfTrialType)
  
  # Group by test type and session, then summarize to calculate accuracy rate
  accuracy_rates <- data %>%
    filter(!is.na(responseCorrectChoice), 
           !is.na(assinfTrialType), !is.na(session),
           assinfTrialType != "", session != "", responseCorrectChoice !="") %>%
    mutate(responseCorrectChoice = as.logical(responseCorrectChoice)) %>%
    group_by(assinfTrialType, session) %>%
    summarise(AccuracyRate = mean(responseCorrectChoice, na.rm = TRUE) * 100) %>%
    ungroup()

  # If you want separate columns for each session as well, you can pivot wider twice, first on session then on test type. This will create column names like 'AB_0', 'AB_1', 'BC_0', 'BC_1', 'AC_0', 'AC_1' etc.
  result_df <- accuracy_rates %>%
    unite("test_session", assinfTrialType, session, sep = "_") %>%
    spread(key = test_session, value = AccuracyRate)

  # Rename columns to include prefix "assinf_choice_"
  new_names <- paste0("assinf_choice_", names(result_df))
  names(result_df) <- new_names
  
  # Define the desired order of your test_sessions
  desired_order <- c("AB_0", "AB_1", "BC_0", "BC_1", "AC_0", "AC_1")

  # Arrange the columns by the desired order
  result_df <- result_df[,paste0("assinf_choice_", desired_order)]
}

# Apply the function to each file and combine the results
assinf_choice_acc <- map_df(csv_files, process_assinf_choice_acc)

# Combine the data frames column-wise
GenRepRec_clean <- cbind(GenRepRec_clean, assinf_choice_acc)
```

### associative inference source trials

Calculate the accuracy rate of associative inference source trials (by trial type and session) across all csv files.

```{r}
# Function to process each CSV file
process_assinf_source_acc <- function(file_path) {
  # Read the data
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Convert relevant columns to factor
  data$assinfTrialType <- as.factor(data$assinfTrialType)
  
  # Group by test type and session, then summarize to calculate accuracy rate
  accuracy_rates <- data %>%
    filter(!is.na(responseCorrectSource), 
           !is.na(assinfTrialType), !is.na(session),
           assinfTrialType != "", session != "", responseCorrectSource !="") %>%
    mutate(responseCorrectSource = as.logical(responseCorrectSource)) %>%
    group_by(assinfTrialType, session) %>%
    summarise(AccuracyRate = mean(responseCorrectSource, na.rm = TRUE) * 100) %>%
    ungroup()

  # If you want separate columns for each session as well, you can pivot wider twice, first on session then on test type. This will create column names like 'AB_0', 'AB_1', 'BC_0', 'BC_1', 'AC_0', 'AC_1' etc.
  result_df <- accuracy_rates %>%
    unite("test_session", assinfTrialType, session, sep = "_") %>%
    spread(key = test_session, value = AccuracyRate)
  
  # Rename columns to include prefix "assinf_source_"
  new_names <- paste0("assinf_source_", names(result_df))
  names(result_df) <- new_names
  
  # Define the desired order of your test_sessions
  desired_order <- c("AB_0", "AB_1", "BC_0", "BC_1", "AC_0", "AC_1")

  # Arrange the columns by the desired order 
  result_df <- result_df[,paste0("assinf_source_", desired_order)]
}

# Apply the function to each file and combine the results
assinf_source_acc <- map_df(csv_files, process_assinf_source_acc)

# Combine the data frames column-wise
GenRepRec_clean <- cbind(GenRepRec_clean, assinf_source_acc)
```

### semantic representation trials

Calculate the accuracy rate of semantic representation trials (by session) across all csv files.

```{r}
# Function to process each CSV file
process_semrep_acc <- function(file_path) {
  # Read the data
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Group by session, then summarize to calculate accuracy rate
  result_df <- data %>%
    filter(!is.na(responseCorrectSem), !is.na(session), session != "", responseCorrectSem !="") %>%
    mutate(responseCorrectSem = as.logical(responseCorrectSem)) %>%
    group_by(session) %>%
    summarise(AccuracyRate = mean(responseCorrectSem, na.rm = TRUE) * 100) %>%
    ungroup() %>%
    pivot_wider(names_from = session, values_from = AccuracyRate)

  # Rename columns to include prefix "semrep_"
  new_names <- paste0("semrep_", names(result_df))
  names(result_df) <- new_names
  
  return(result_df)
}

# Apply the function to each file and combine the results
semrep_acc <- map_df(csv_files, process_semrep_acc)

# Combine the data frames column-wise
GenRepRec_clean <- cbind(GenRepRec_clean, semrep_acc)
```

### visual representation trials

Calculate the accuracy rate of visual representation trials (by session) across all csv files.

```{r}
# Function to process each CSV file
process_visrep_acc <- function(file_path) {
  # Read the data
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Group by session, then summarize to calculate accuracy rate
  result_df <- data %>%
    filter(!is.na(responseCorrectVis), !is.na(session), session != "", responseCorrectVis !="") %>%
    mutate(responseCorrectVis = as.logical(responseCorrectVis)) %>%
    group_by(session) %>%
    summarise(AccuracyRate = mean(responseCorrectVis, na.rm = TRUE) * 100) %>%
    ungroup() %>%
    pivot_wider(names_from = session, values_from = AccuracyRate)

  # Rename columns to include prefix "visrep_"
  new_names <- paste0("visrep_", names(result_df))
  names(result_df) <- new_names
  
  return(result_df)
}

# Apply the function to each file and combine the results
visrep_acc <- map_df(csv_files, process_visrep_acc)

# Combine the data frames column-wise
GenRepRec_clean <- cbind(GenRepRec_clean, visrep_acc)
```

### recognition memory trials

Calculate the accuracy rate of recognition memory trials (by trial type and session) across all csv files.

```{r}
# Function to process each CSV file
process_rec_acc <- function(file_path) {
  # Read the data
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Convert relevant columns to factor
  data$recTrialType <- as.factor(data$recTrialType)
  
  # Group by test type and session, then summarize to calculate accuracy rate
  accuracy_rates <- data %>%
    filter(!is.na(responseCorrectRec), 
           !is.na(recTrialType), !is.na(session),
           recTrialType != "", session != "", responseCorrectRec !="") %>%
    mutate(responseCorrectRec = as.logical(responseCorrectRec)) %>%
    group_by(recTrialType, session) %>%
    summarise(AccuracyRate = mean(responseCorrectRec, na.rm = TRUE) * 100) %>%
    ungroup()

  # If you want separate columns for each session as well, you can pivot wider twice, first on session then on test type. 
  result_df <- accuracy_rates %>%
    unite("test_session", recTrialType, session, sep = "_") %>%
    spread(key = test_session, value = AccuracyRate)
  
  # Rename columns to include prefix "rec_"
  new_names <- paste0("rec_", names(result_df))
  names(result_df) <- new_names
  
  # Define the desired order of your test_sessions
  desired_order <- c("Old_0", "Old_1", "Lure_0", "Lure_1", "New_0", "New_1")

  # Arrange the columns by the desired order 
  result_df <- result_df[,paste0("rec_", desired_order)]
}

# Apply the function to each file and combine the results
rec_acc <- map_df(csv_files, process_rec_acc)

# Combine the data frames column-wise
GenRepRec_clean <- cbind(GenRepRec_clean, rec_acc)

# Define the weights
weights_0 <- c(rec_Old_0 = 32/95, rec_Lure_0 = 31/95, rec_New_0 = 32/95)
weights_1 <- c(rec_Old_1 = 31/95, rec_Lure_1 = 32/95, rec_New_1 = 32/95)

# Calculate the weighted mean for each row
GenRepRec_clean$rec_Overall_0 <- (GenRepRec_clean$rec_Old_0 * weights_0[1] + GenRepRec_clean$rec_Lure_0 * weights_0[2] + GenRepRec_clean$rec_New_0 * weights_0[3]) 
GenRepRec_clean$rec_Overall_1 <- (GenRepRec_clean$rec_Old_1 * weights_1[1] + GenRepRec_clean$rec_Lure_1 * weights_1[2] + GenRepRec_clean$rec_New_1 * weights_1[3]) 
```

## Restructuring the survey and demographics data

Extract the JSON string and parse it into a new data frame for survey questions.

```{r}
# Function to process a single file
process_survey_data <- function(file_name) {
    # Read the CSV file
    df <- read.csv(file_name, stringsAsFactors = FALSE)
    
    # Extract the JSON string from the third-to-last row of the 'response' column
    json_str <- df$response[nrow(df) - 2]
    
    # Parse the JSON string into a data frame
    parsed_data <- fromJSON(json_str)
    
    return(as.data.frame(parsed_data))
}

# Process each file and combine results
results_survey <- map_df(csv_files, process_survey_data)

# Rename the columns from 'Q0'-'Q8' to 'Q1'-'Q9'
colnames(results_survey) <- paste0('Q', 1:9)

# Combine the data frames column-wise
GenRepRec_clean <- cbind(GenRepRec_clean, results_survey)
```

Extract the JSON string and parse it into a new data frame for demographics information.

```{r}
# Function to process a single file
process_demographics <- function(file_name) {
    # Read the CSV file
    df <- read.csv(file_name, stringsAsFactors = FALSE)
    
    # Extract the JSON string from the second-to-last row of the 'response' column
    json_str <- df$response[nrow(df) - 1]
    
    # Parse the JSON string into a data frame
    parsed_data <- fromJSON(json_str)
    
    return(as.data.frame(parsed_data))
}

# Process each file and combine results
results_demographics <- map_df(csv_files, process_demographics)

# Combine the data frames column-wise
GenRepRec_clean <- cbind(GenRepRec_clean, results_demographics)
```

## Calculating AC trials accuracy after removing triads with incorrect AB and BC trials

Choice trials:

```{r}
process_directcorrect_AC_choice_acc <- function(file_path) {
  # Read the data
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Convert relevant columns to string and lowercase for consistency
  data$responseCorrectChoice <- tolower(as.character(data$responseCorrectChoice))
  data$assinfTrialType <- as.character(data$assinfTrialType)
  
  # Create a list of triadIDs for all AB and BC trials where the response was incorrect
  incorrect_ab_bc_triadIDs <- data %>%
    filter(assinfTrialType %in% c('AB', 'BC'), responseCorrectChoice == 'false') %>%
    pull(triadID) # Extracting the triadID directly
  
  # Filter for AC trials
  ac_trials <- data %>%
    filter(assinfTrialType == 'AC' & (responseCorrectChoice == "false" | responseCorrectChoice == "true"))

  # Filter out AC trials where triadID matches any of the incorrect AB/BC triadIDs
  correct_ac_trials <- ac_trials %>%
    filter(!(triadID %in% incorrect_ab_bc_triadIDs))
  
  # Calculate accuracy rate by session, using the filtered AC trials
  session_accuracy <- correct_ac_trials %>%
    mutate(responseCorrectChoice = responseCorrectChoice == "true") %>%
    group_by(session) %>%
    summarise(
      DirectCorrect_AC_Choice_Acc = mean(responseCorrectChoice, na.rm = TRUE) * 100,
      NumberOf_DirectCorrect_AC = n()
    ) %>%
    ungroup() %>%
    pivot_longer(-session, names_to = "Metric", values_to = "Value") %>%
    unite("SessionMetric", Metric, session, sep = "_") %>%
    pivot_wider(names_from = "SessionMetric", values_from = "Value")
  
  return(session_accuracy)
}

# Apply the function to each file and combine the results
directcorrect_AC_choice_acc <- map_df(csv_files, process_directcorrect_AC_choice_acc)

# Combine the data frames column-wise
GenRepRec_clean <- cbind(GenRepRec_clean, directcorrect_AC_choice_acc)
```

Source trials:

```{r}
process_directcorrect_AC_source_acc <- function(file_path) {
  # Read the data
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Convert relevant columns to string and lowercase for consistency
  data$responseCorrectSource <- tolower(as.character(data$responseCorrectSource))
  data$assinfTrialType <- as.character(data$assinfTrialType)
  
  # Create a list of triadIDs for all AB and BC trials where the response was incorrect
  incorrect_ab_bc_triadIDs <- data %>%
    filter(assinfTrialType %in% c('AB', 'BC'), responseCorrectSource == 'false') %>%
    pull(triadID) # Extracting the triadID directly
  
  # Filter for AC trials
  ac_trials <- data %>%
    filter(assinfTrialType == 'AC' & (responseCorrectSource == "false" | responseCorrectSource == "true"))

  # Filter out AC trials where triadID matches any of the incorrect AB/BC triadIDs
  correct_ac_trials <- ac_trials %>%
    filter(!(triadID %in% incorrect_ab_bc_triadIDs))
  
  # Calculate accuracy rate by session, using the filtered AC trials
  session_accuracy <- correct_ac_trials %>%
    mutate(responseCorrectSource = responseCorrectSource == "true") %>%
    group_by(session) %>%
    summarise(
      DirectCorrect_AC_Source_Acc = mean(responseCorrectSource, na.rm = TRUE) * 100,
      NumberOf_DirectCorrect_AC_Source = n()
    ) %>%
    ungroup() %>%
    pivot_longer(-session, names_to = "Metric", values_to = "Value") %>%
    unite("SessionMetric", Metric, session, sep = "_") %>%
    pivot_wider(names_from = "SessionMetric", values_from = "Value")
  
  return(session_accuracy)
}

# Apply the function to each file and combine the results
directcorrect_AC_source_acc <- map_df(csv_files, process_directcorrect_AC_source_acc)

# Combine the data frames column-wise
GenRepRec_clean <- cbind(GenRepRec_clean, directcorrect_AC_source_acc)
```

## Calculating associative inference choice/source accuracy and reaction time by cue image's position across different trial types

```{r}
process_assinf_bycue <- function(file_path, test_col) {
  # Read the data
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Ensure column names are correctly referenced as symbols
  test_sym <- rlang::sym(test_col)
  
  # Pre-process data
  data <- data %>%
    filter(!is.na(.data[[test_sym]]), !is.na(assinfTrialType), !is.na(cuePosition),
           assinfTrialType != "", cuePosition != "", .data[[test_sym]] != "") %>%
    mutate(test = as.logical(.data[[test_sym]]),
           rt = as.numeric(rt))
  
  # Calculate accuracy and count for each trial type and cue position
  accuracy_data <- data %>%
    group_by(assinfTrialType, cuePosition, session) %>%
    summarise(accuracy = mean(test, na.rm = TRUE) * 100,
              rt = mean(rt, na.rm = TRUE),
              count = n(), 
              .groups = 'drop')
  
  # Combine accuracy and count into one column with a key
  accuracy_long <- accuracy_data %>%
    pivot_longer(c(accuracy, rt, count), names_to = "metric", values_to = "value")

  # Pivot to wide format with combined metric names
  accuracy_wide <- accuracy_long %>%
    unite("condition_metric", assinfTrialType, cuePosition, metric, session, sep = "_") %>%
    pivot_wider(names_from = condition_metric, values_from = value)

  return(accuracy_wide)
}

# Apply the function to each file and combine the results for Choice tests
assinf_choice_bycue <- map_df(csv_files, ~process_assinf_bycue(.x, "responseCorrectChoice"))
colnames(assinf_choice_bycue) <- paste0("choice_", colnames(assinf_choice_bycue))

# Apply the function to each file and combine the results for Source tests
assinf_source_bycue <- map_df(csv_files, ~process_assinf_bycue(.x, "responseCorrectSource"))
colnames(assinf_source_bycue) <- paste0("source_", colnames(assinf_source_bycue))

# Combine the data frames column-wise
GenRepRec_clean <- cbind(GenRepRec_clean, assinf_choice_bycue)
GenRepRec_clean <- cbind(GenRepRec_clean, assinf_source_bycue)
```

## Calculating mean reaction time for the associative inference test

### associative inference choice trials

```{r}
# Function to process each CSV file
process_assinf_choice_rt <- function(file_path) {
  # Read the data
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Convert relevant columns to string
  data$assinfTrialType <- as.factor(data$assinfTrialType)
  
  # Group by test type and session, then summarize to calculate accuracy rate
  mean_rts <- data %>%
    filter(!is.na(responseCorrectChoice), 
           !is.na(assinfTrialType), !is.na(session),
           assinfTrialType != "", session != "", responseCorrectChoice !="") %>%
    mutate(rt = as.numeric(rt)) %>%
    group_by(assinfTrialType, session) %>%
    summarise(rt = mean(rt, na.rm = TRUE)) %>%
    ungroup()

  # If you want separate columns for each session as well, you can pivot wider twice, first on session then on test type. This will create column names like 'AB_0', 'AB_1', 'BC_0', 'BC_1', 'AC_0', 'AC_1' etc.
  result_df <- mean_rts %>%
    unite("test_session", assinfTrialType, session, sep = "_") %>%
    spread(key = test_session, value = rt)

  # Rename columns to include prefix "assinf_choice_"
  new_names <- paste0("rt_assinf_choice_", names(result_df))
  names(result_df) <- new_names
  
  # Define the desired order of your test_sessions
  desired_order <- c("AB_0", "AB_1", "BC_0", "BC_1", "AC_0", "AC_1")

  # Arrange the columns by the desired order
  result_df <- result_df[,paste0("rt_assinf_choice_", desired_order)]
}

# Apply the function to each file and combine the results
assinf_choice_rt <- map_df(csv_files, process_assinf_choice_rt)

# Combine the data frames column-wise
GenRepRec_clean <- cbind(GenRepRec_clean, assinf_choice_rt)
```

### associative inference source trials

```{r}
# Function to process each CSV file
process_assinf_source_rt <- function(file_path) {
  # Read the data
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Convert relevant columns to string
  data$assinfTrialType <- as.factor(data$assinfTrialType)
  
  # Group by test type and session, then summarize to calculate accuracy rate
  mean_rts <- data %>%
    filter(!is.na(responseCorrectSource), 
           !is.na(assinfTrialType), !is.na(session),
           assinfTrialType != "", session != "", responseCorrectSource !="") %>%
    mutate(rt = as.numeric(rt)) %>%
    group_by(assinfTrialType, session) %>%
    summarise(rt = mean(rt, na.rm = TRUE)) %>%
    ungroup()

  # If you want separate columns for each session as well, you can pivot wider twice, first on session then on test type. This will create column names like 'AB_0', 'AB_1', 'BC_0', 'BC_1', 'AC_0', 'AC_1' etc.
  result_df <- mean_rts %>%
    unite("test_session", assinfTrialType, session, sep = "_") %>%
    spread(key = test_session, value = rt)

  # Rename columns to include prefix "assinf_choice_"
  new_names <- paste0("rt_assinf_source_", names(result_df))
  names(result_df) <- new_names
  
  # Define the desired order of your test_sessions
  desired_order <- c("AB_0", "AB_1", "BC_0", "BC_1", "AC_0", "AC_1")

  # Arrange the columns by the desired order
  result_df <- result_df[,paste0("rt_assinf_source_", desired_order)]
}

# Apply the function to each file and combine the results
assinf_source_rt <- map_df(csv_files, process_assinf_source_rt)

# Combine the data frames column-wise
GenRepRec_clean <- cbind(GenRepRec_clean, assinf_source_rt)
```

## Convert back to the CSV file

```{r}
# Write the updated data frame back to the CSV file
write.csv(GenRepRec_clean, 'GenRepRec_clean_TwoSessionsVer.csv', row.names = FALSE)
```
